Assessing the reliability of an online behavioural laterality battery: a pre-registered study.

### 2020-11-08 online_laterality_reliability.Rmd
include some responses to reviewers

---
title: 'Assessing the reliability of an online behavioural laterality battery: a pre-registered study.'
output:
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

<!--- updated 20/7/20 to change 'select' to 'dplyr::select' to avoid conflicts-->
<!--- updated 3/8/20 to add ZW feedback--->
Adam J. Parker*

Zoe V. J. Woodhead

Paul A. Thompson

Dorothy V. M. Bishop

@@

Behavioural laterality tasks have long been used to establish hemispheric dominance for certain cognitive functions. Results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018). Here we review how each paradigm has been implemented to study language lateralisation and what is known about reliability and validity.

#### Visual half-field paradigm

In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus. The visual pathways from the left and right sides of space to the visual cortex are crossed for non-foveal viewing, and so the right visual-field (RVF) projects directly to the left hemisphere and the left visual-field (LVF) to the right hemisphere. The corpus callosum allows for transfer between hemispheres, but the rationale of the method is that there will be a processing advantage for language stimuli that are directly projected to the language-dominant hemisphere. This method has been used since the 1950s (Mishkin & Forgays, 1952). Bourne (2006) reviewed the literature, which consistently shows advantageous processing for letter strings presented in the RVF, relative to the LVF, in most people.

@@

In recent years it has become clear that the reliability and validity of visual half-field tasks to identify the language-dominant hemisphere is largely dependent on several methodological aspects. For example, Hunter and Brysbaert (2008) noted stronger and more stable visual field differences when stimuli were presented bilaterally relative to those where only one stimulus was presented either in the LVF or in the RVF (Boles, 1987, 1990, 1994; see also Iacoboni & Zaidel, 1996). Hunter and Brysbaert argued that this is the result of stimuli competing during bilateral presentation, enhancing the advantage to the stimulus presented in the language dominant hemisphere. Reliability and validity also depend on there being a sufficient number of trials and brief stimulus presentation to prevent eye movements. When accounting for these methodological aspects, the validity of visual half-field tasks seems promising. For example, Hunter and Brysbaert (2008) used two visual half-field tasks, word naming (192 trials) and picture naming (160 trials) in which stimuli were bilaterally presented and masked at offset. The two tasks correlated strongly (*r*= .8, 95% CI .598-.956) in a sample of 26 left-handers. Encouragingly, the LIs from these visual half-field tasks also correlated positively with laterality from functional Magnetic Resonance Imaging (fMRI) on a word generation task in 10 left-handers. However, the small sample size meant that the confidence intervals around these estimates were very wide: word naming, *r*= .63 (95% CI = .001-.902), picture naming, *r*= .77 (95% CI = .272-.943).

#### Dichotic listening

@@

#### Chimeric faces

@@

With regards to visual half-field tasks for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Weekes, Capetillo-Cunliffe, Rayman, Iacoboni, & Zaidel, 1999). However, a relatively well-powered study involving 100 participants, conducted by Willemin et al. (2016), reported no hint of any effect of handedness, when coded as left and right, or when divided into left (N = 19), mixed (N = 24) and right-handed (N = 57) groups, on the visual field advantage during a lexical decision task. This result illustrates the recurring challenge of interpreting null results in this field: do they reflect a true lack of relationship, poor test reliability, inadequate power to detect small effects, or specific methodological factors? We may note that Willemin et al. asked participants to respond via manual response, whereas other studies have involved an active speech production component, which could influence task lateralisation.

@@

From the above-reviewed studies, it appears that handedness is related to laterality across a range of tasks (for further discussion see Carey & Johnstone, 2014). This relationship points towards a general laterality factor which influences lateralisation across tasks and behaviours, as predicted by several theoretical accounts (Behrmann & Plaut, 2015; Hellige, 1993; Kosslyn, 1987). In a study of left-handers using fMRI with five tasks, Gerrits, Verhelst, and Vingerhoets (2020) found that the just under half the participants showed either standard laterality (left-lateralised for language and praxis, right-lateralised for face memory, line-bisection and vocal emotion) or a complete reversal of this pattern, providing clear evidence against the idea that these functions are lateralised independently. It is also clear that this association is far from perfect, (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018). Our ultimate goal is to unpick whether dissociations between laterality indices are evidence of multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.

@@

The principal goal of the current study was to evaluate the reliability and validity of an online behavioural laterality battery. Specifically, we examined whether it is possible to observe the usual laterality biases at the population level in tests of language and emotion processing and whether it is possible to achieve test-retest reliability of *r*= .65 or above using online behaviour research methods. In addition, for tasks where satisfactory reliability was achieved, we considered how many trials were needed to achieve this. The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening task, a chimeric faces visual half-field task, and a novel finger tapping task.<!--- ZW: Would it be worth having a section on measurement of manual preference?; AP: As this is exploratory I don't want to add too much about this---> These behavioural tasks were completed twice on two different occasions.

@@

Our final prediction (hypothesis 3) related to the inter-relationships between laterality indices. With online testing, there is the opportunity to test inter-relationships between laterality indices with an adequately powered sample using measures of known reliability. Prior studies have been inconclusive because it has been difficult to gather large amounts of data to establish test-retest reliability. For instance, van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks. However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of the individual measures. In the preregistration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (model A) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model B), or a base model where each task is independently lateralised (model C). Given our previous findings using fTCD to study dissociations in language laterality (Woodhead, Bradshaw, Wilson, Thompson, & Bishop, 2019), we further predicted that model fit may differ for right- and left-handers, with more dissociation between tasks in the latter group.

@@

In practice, our final sample exceeded 300, because we discarded participants from individual tasks for failing to meet prespecified criteria (see below). In total, 441 native English speakers were initially recruited through Prolific Academic (www.prolific.co). Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g. head injury, brain tumour, stroke) or significant hearing loss or impairment.

@@

Adapted versions of the Miles test (Miles, 1929) and the Porta test (Porac & Coren, 1976) were used to determine each participant’s eye dominance in central gaze (i.e., when looking straight ahead). Each test classified participants as being either left or right eye dominant. In practice, the two tests agreed perfectly, so we report here just the Miles test.

@@

The rhyme decision task was developed for this project as a novel visual half-field technique designed to engage lateralised language processes. On each trial, participants were asked to decide whether a foveally presented letter string rhymed with the name of an image shown in either the LVF or RVF. With online presentation, a particular concern is a potential for eye movements away from fixation. We relied on the fact that participants had to focus initially on the central written word to do the task, with pictures subsequently displayed with brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982). Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972; but see also Bourne, 2006). Pictures of familiar objects were presented for rhyme matching to the written word, to ensure that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing.

@@

![Figure 1. Schematic illustration of a trial on the rhyme decision task. In the example, the word “bite” is presented along with the two images “kite” and “door”. When making the correct judgement, participants would respond with by pressing ‘S’ to indicate that the rhyme was on the left.](Picture1.png)

@@

Each trial began with a fixation cross presented centrally for 800 ms. Participants were instructed to keep their eyes on the central fixation point. Then a foveally presented word would appear for 200 ms. The bilaterally presented images were then presented for 150 ms. At stimulus offset, participants indicated whether the name of the image in the LVF or RVF rhymed with the foveally presented word. Participants were instructed to press ‘S’ if the written word rhymed with the image in the LVF, and ‘K’ if it rhymed with the image in the RVF. If participants did not detect a rhyme, they were instructed to press the space bar. The response triggered the next trial. Accuracy and reaction times (RT) were recorded.

@@

As the dichotic listening task demands the use of stereo-headphones, participants were screened at the start of each session. Screening consisted of two tasks. The first involved participants judging which of three pure tones was quietest, and is described by Woods, Siegel, Traer, and McDermott (2017) in full detail. This sound discrimination task was designed to be easy to achieve when using headphones but difficult over loudspeakers due to phase-cancellation. Participants were removed from the study if they failed to correctly discriminate the quietest sound on at least five out of six trials, indicating that they were unlikely to be using headphones. The second check involved participants hearing a sound presented in only one of the two channels. Participants were removed if they failed to correctly identify which channel the sound was played on five out of six trials, indicating that they were not using stereo-headphones.

Participants completed four blocks of the dichotic listening task. In each block, participants heard each of the 36 stimuli once. Across the four blocks, they responded to 144 stimuli. At the start of each trial, participants saw a fixation cross for 250 ms, after which a stimulus was played. After hearing each stimulus, participants were instructed to report the syllable they heard or if it seemed like they heard two different sounds, the one they heard most clearly by clicking a button corresponding to one of the six syllables. The response triggered the start of the next trial. Accuracy and reaction times (RT) were recorded.

@@

![Figure 2. An example stimulus from the chimeric faces task. In the example, “happiness” is shown on the left and “sadness” on the right. A participant who is right lateralised for emotional processing would be more likely to perceive the emotion in the left hemifield and would, therefore, respond “happiness”.](chim_example.png)

@@

Participants completed four blocks of the chimeric faces task. In each block, participants viewed each stimulus twice. Thus, they viewed 32 stimuli per block and 128 in total. At the start of each trial, participants fixated a cross for 1000 ms. A chimeric face was then displayed for 400 ms. Participants then reported the emotion seen in the face by clicking a button corresponding to one of the four emotions. The response triggered the start of the next trial. Accuracy and reaction times (RT) were recorded.

@@

The current study implemented a test-retest, within-subjects design across two sessions. Each session was three to 10 days apart. During session 1, participants completed the full battery of tasks. Self -report measures were completed first, followed by the behavioural laterality tasks. During session 2, participants completed only the behavioural laterality tasks. Hence, each participant provided data for the rhyme decision, dichotic listening, chimeric faces, and finger tapping tasks twice. Each session lasted approximately 45 minutes and 35 minutes respectively.

When completing the behavioural tasks participants were instructed to sit approximately 50 cm from the screen. Gorilla’s scaling tool was utilised such that stimuli size and eccentricity was maintained across tasks and participants. Stimuli from each task were repeated across four blocks. Within each block, participants completed 70 trials of the rhyme detection task (R); 36 trials of the dichotic listening task (D); 32 trials of the chimeric face task (C); and 2 trials of the finger tapping task (F) for both their dominant and non-dominant hand. The order of tasks with each block was determined as follows RDCF, DCFR, CFRD, FRDC, and the order of tasks remained consistent across blocks for each participant. For each task, trial order was randomised, except for the finger tapping task where participants responded with their dominant hand followed by their non-dominant hand.

@@

For each task, we adopted the following data cleaning procedures. First, for all tasks, we removed trials in which response times were shorter than 200 ms. Response times were log-transformed to reduce skew, as specified in our pre-registration. Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)). Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution. The Hoaglin and Iglewicz procedure was applied to each participant's data separately. After applying each data trimming procedure, participants left with less than 80% of trials were removed.

@@

Foot and eye preference characteristics of the sample after exclusions as noted above are shown in Table 1. As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we report data only from the Miles test.

@@

}
```

>>>>

---
title: 'Assessing the reliability of an online behavioural laterality battery: a pre-registered study.'
output:
  html_notebook: default
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r font, echo- FALSE, include= FALSE}
library(officedown)
library(officer)
ft <- fp_text(font.size = 12, bold = FALSE, color = "#C32900", font.family = "Cambria")

colFmt = function(x,color){
  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

<!--- updated 20/7/20 to change 'select' to 'dplyr::select' to avoid conflicts-->

<!--- updated 3/8/20 to add ZW feedback--->
\begin{center}
Adam J. Parker*, Zoe V. J. Woodhead, Paul A. Thompson, Dorothy V. M. Bishop
\end{center}

@@

Behavioural laterality tasks have long been used to establish hemispheric dominance for certain cognitive functions. Results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018). `r colFmt("By contrast, visual half-field paradigms indicate that the right hemisphere is involved in the processing of faces, global feature processing, and visuo-spatial attention (Brederoo, Nieuwenstein, Cornelissen,",'red')` & `r colFmt("Lorist, 2019). Below",'red')` we review how each paradigm has been implemented to study language lateralisation `r colFmt("and the detection of emotion in faces", 'red')` and what is known about reliability and validity.

#### Lateralised visual word processing

`r colFmt("Lateralised visual word processing has been studies using the visual half-field paradigm.", 'red')`In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus. The visual pathways from the left and right sides of space to the visual cortex are crossed for non-foveal viewing, and so the right visual-field (RVF) projects directly to the left hemisphere and the left visual-field (LVF) to the right hemisphere. The corpus callosum allows for transfer between hemispheres, but the rationale of the method is that there will be a processing advantage for language stimuli that are directly projected to the language-dominant hemisphere. This method has been used since the 1950s (Mishkin & Forgays, 1952). Bourne (2006) reviewed the literature, which consistently shows advantageous processing for letter strings presented in the RVF, relative to the LVF, in most people.

@@

In recent years it has become clear that the reliability and validity of visual half-field tasks to identify the language-dominant hemisphere is largely dependent on several methodological aspects. For example, Hunter and Brysbaert (2008) noted stronger and more stable visual field differences when stimuli were presented bilaterally relative to those where only one stimulus was presented either in the LVF or in the RVF (Boles, 1987, 1990, 1994; see also Iacoboni & Zaidel, 1996). Hunter and Brysbaert argued that this is the result of stimuli competing during bilateral presentation, enhancing the advantage to the stimulus presented in the language dominant hemisphere. Reliability and validity also depend on there being a sufficient number of trials and brief stimulus presentation to prevent eye movements. When accounting for these methodological aspects, the validity of visual half-field tasks seems promising. For example, Hunter and Brysbaert (2008) used two visual half-field tasks, word naming (192 trials) and picture naming (160 trials) in which stimuli were bilaterally presented and masked at offset. The two tasks correlated strongly (*r*= .8, 95% CI .598-.956) in a sample of 26 left-handers. Encouragingly, the LIs from these visual half-field tasks also correlated positively with laterality from functional Magnetic Resonance Imaging (fMRI) on a word generation task in 10 left-handers. However, the small sample size meant that the confidence intervals around these estimates were very wide: word naming, *r*= .63 (95% CI = .001-.902), picture naming, *r*= .77 (95% CI = .272-.943).`r colFmt("More recently, however, Van der Haegen, Cai, Seurinck, and Brysbaert (2011) reported highly similar findings in a somewhat larger sample of 50 participants. Again, participants completed behavioural word and picture naming visual half-field tasks and then a word generation task under fMRI. It is reassuring that the LIs for the behavioural measures were correlated with the fMRI LI",'red')`: `r colFmt("word naming,",'red')` *r*`r colFmt("= .64, picture naming, ",'red')` *r* `r colFmt("= .65. Van der Haegen et al. also report a set of analyses involving 250 participants who completed the behavioural visual half-field tasks while their eye movements were monitored. Two sets of LIs were calculated",'red')`; `r colFmt("one controlling for fixation error and one that did not. Importantly, the LIs which did and did not control for fixation error were highly correlated for word naming, ", 'red')`*r*`r colFmt("= .94, and picture naming, ",'red')`*r*`r colFmt("=.98, indicating that the inclusion of eye movement control adds little additional value when a visual half-field task employs mechanisms to encoruage central fixation and bilateral presentation.",'red')`

`r colFmt("A promising line of work has emerged which further highlights the potenitial for well designed visual half-field tasks to assess language lateralisation. First introduced by Willemin et al. (2016), the translingual lexical decision task involves participants viewing letter stings either side of fixation and judging whether the left, right, or neither of the letter strings was a valid word. With 100 participants completing 256 trials, Willemin et al. reported a RVF advatage at the population level. Since then this effect has been replicated in a large sample 496 participants cvering six international languages, with evidence of a RVF for each language studied. Not only does this task show validity, it seems reliable given Brederoo et al.'s (2020) obervsation that the task yields good split-half reliability when tested with 122 participants,",'red')` *r*`r colFmt("= 0.86 for error rates and", 'red')` *r*`r colFmt("= 0.80 for reaction times.",'red')`

#### Laterlised auditory processing

@@

`r colFmt("To date, little work has been conducted online with visual stimuli. By contrast, the dichotic listening paradigm has been implemented online with huge success. Bless et al. (2013) developed a mobile device version of the consonant-voewl dichotic listening paradigm, with 36 simultaneous sound pairs. Under laboratory conditions, with a sample of 76 (33 Norwegian, 43 Austrialian) participants, they reported a test-retest reliability of *r*= 0.78 when using the mobile app. Furthermore, both subgroups of participants showed clear right ear advantages at the population level. In a subsequent large-scale study, Bless et al. (2015) administered the same application to a sample of over 4,000 participants with more than 60 different language backgrounds. Again Bless et al. (2015) reported a clear right ear advanatge at the population-level. They then used the large dataset to examine cultural/lingusitic effects on the degree on lateralisation; reporting that the degree of lateralization appears to be modulated by linguistic background. Together these studies demonstrate the potenital for online laterality research to generate large amounts of data while maintaining a tolerable degree of reliability.",'red')`

#### Lateralised face processing

@@

With regards to visual half-field tasks for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Weekes, Capetillo-Cunliffe, Rayman, Iacoboni, & Zaidel, 1999). However, a relatively well-powered study involving 100 participants, conducted by Willemin et al. (2016), reported no hint of any effect of handedness, when coded as left and right, or when divided into left (N = 19), mixed (N = 24) and right-handed (N = 57) groups, on the visual field advantage during a lexical decision task. This result illustrates the recurring challenge of interpreting null results in this field: do they reflect a true lack of relationship, inadequate power to detect small effects, or specific methodological factors? `r colFmt("At current, we can rule out poor reliability as a cause for this null result as Brederoo et al. (2020) reported good split-half reliability for this task.",'red')` We may note that Willemin et al. asked participants to respond via manual response, whereas other studies have involved an active speech production component, which could influence task lateralisation `r colFmt("and may have contributed to the null effect of handedness.", 'red')`

@@

From the above-reviewed studies, it appears that handedness is related to laterality across a range of tasks (for further discussion see Carey & Johnstone, 2014). This relationship points towards a general laterality factor which influences lateralisation across tasks and behaviours, as predicted by several theoretical accounts (Behrmann & Plaut, 2015; Hellige, 1993; Kosslyn, 1987). `r colFmt("For example, the causal complementarity principle proposes that the lateralisation of processes involving the same hemisphere will be lateralised to the homologue in the opposite hemisphere. With reference to language processing and face emotionality processing, it would be predicted that language processing would be left lateralised and face processing would be right lateralised, and an increase of one co-occurs with an increase of another. Indeed, several studies have reported such patterns. In support, Brederoo et al. (2020) and Dundas, Plaut, and Behrmann (2015) have reported significant correlations between LIs for word and face processing. However, that does not mean that all studies provide support for causal complementarity",'red')`. In a study of left-handers using fMRI with five tasks, Gerrits, Verhelst, and Vingerhoets (2020) found that the just under half the participants showed either standard laterality (left-lateralised for language and praxis, right-lateralised for face memory, line-bisection and vocal emotion) or a complete reversal of this pattern, providing clear evidence against the idea that these functions are lateralised independently. It is also clear that this association is far from perfect (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018). Our ultimate goal is to unpick whether dissociations between laterality indices are evidence of multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.

`r colFmt("So far the argument has been made that lateralisation across domains do not correlate. However, emerging lines of evidence have suggested that even within domains there is independent lateralisation across similar tasks. Woodhead, Bradshaw, Wilson, Thompson, and Bishop (2019) reported data for 37 participants engaged in six language tasks under transcranial Doppler sonography (fTCD) and examined the factorial structure of lateralisation across tasks using structural equation modeling. They pitted two models against each other. The first assumed that language laterality is a unidimensional trait and a single factor was used to model covariance. The second assumed that laterality on one set of tasks may vary independently with that one another. This model assumed that two factors underlie the pattern of covariance. The results supported a two factor model at the population level. However, further examination of the data revealed that this was driven by seven left-handers. For the majority of right-handed participants a single factor model was preferred. Of course this pattern of results may have resulted from the small proportion of left-handers in the current study. Woodhead, Thompson, Karlsson, and Bishop (2020) examined this factor structure in a larger sample of 74 participants with 43 right-handers and 31 left-handers. The data this time indicated that a two factor model was preferred for both right- and left-handers, with the preference for a two factor model being stronger for left-handers. Together, this indicated that not all language tasks show correlated laterality indices. Instead, there was evidence that the tasks covaried in two factors with distinct patterns. For the current work which implements two measures of llanguage laterality, it may be perfectly plausible that we observe a pattern of results which indicate that language lateralisation is not unidimensional and language tasks would fractionate.",'red')`

@@

The principal goal of the current study was to evaluate the reliability and validity of an online behavioural laterality battery. Specifically, we examined whether it is possible to observe the usual laterality biases at the population level in tests of language and emotion processing and whether it is possible to achieve test-retest reliability of *r*= .65 or above using online behaviour research methods. In addition, for tasks where satisfactory reliability was achieved, we considered how many trials were needed to achieve this. The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening task, a chimeric faces visual half-field task, and a novel finger tapping task. These behavioural tasks were completed twice on two different occasions. `r colFmt("Below we justify our task selection.",'red')`.

`r colFmt("When looking to include studies in our online battery we saw three criteria that a task would have to meet for inclusion. First, it would have show good test-retest reliability. Second, it would have to show validity and be lateralised at the population level. Third, it would have to be sensitive to the subtle differences in laterality that have been reported between right- and non-right-handers. Starting with the dichotic listening paradigm, recent endeavors have shown that the task yields reliable results when used online via a mobile app (Bless et al., 2013, 2015). Furthermore, Karlsson et al. (2019) have shown that this paradigm is sensitive to the subtle differences in language asymmetry between right and non-right handers, making it an obvious choice for inclusion. Given that numerous studies looking at the relationship lateralisation on language and non-language tasks have focused on the use of face stimuli (e.g. Brederoo et al., 2020), we chose to include a face task to conduct similar comparisons. Regarding the chimeric faces, Levy et al. (1983) reported good split-half reliability while Karlsson et al. had reported that, like the dichotic listening task, the the chimeric faces task is sensitive to the differences in perceptual asymmetry in right- and non-right-handers. In addition to these two tasks we looked to include a visual half-field task so that we could further assess language lateralisation. While the task reported by Brysbaert and colleagues looked promising, they involve verbal responses which would be difficult to record online with reasonable accuracy. We then explored the use of translingual lexical decision task reported by Hausmann and colleagues. At the point of pre-registering the study there had been no assessment of the tasks reliability. This coupled with the absence of an effect of handedness on reaction times in this task (Willemmin et al., 2016) did not give us faith in using the task. Thus we looked to develop our own task in which would required covert speech.",'red')`

`r colFmt("In addition to the previous mentioned tasks, we sought to develop an online measure of manual performance. This was largely an exploratory effort. Most studies allude to handedness as a unitary phenomenon and classify using a dichotomy of left or right based on participants’ self-reported preference. Despite this, some right-handers will use their left hand more than others (Annett, 1998, 2002). This type of subtle variation is not captured using dichotomous reporting but is captured within some inventories. One such inventory is the Edinburgh handedness inventory (Oldfield, 1971), which requires participants to indicate the strength of hand preference for a range of activities (e.g. throwing, striking a match, using scissors). Inventory measures are not without limitation. It has been argued that respondents may avoid extreme responses on items or solely select extreme responses, thus confounding the measurements (Beaton",'red')` & `r colFmt("Moseley, 1984). A final way to assess handedness is to use performance measures, such as the Purdue Pegboard task (Tiffin",'red')` & `r colFmt("Asher, 1948).  Performance measures can detect subtle differences between each hand which can be placed on a continuum while correlating well with measures of handedness (e.g. Steenhous",'red')` & `r colFmt("Bryden, 1999). One of the most widely used performance tasks is the pegboard task (see Hodgson",'red')` & `r colFmt("Hudson, 2016, for discussion) yet this task is far from employable online given the requirement of pegs and a board. Thus, in addition to examining our hypotheses, we sought to develop and assess the reliability of a novel finger tapping task that could be used to quantify handedness based on performance.",'red')`

@@

Our final prediction (hypothesis 3) related to the inter-relationships between laterality indices. With online testing, there is the opportunity to test inter-relationships between laterality indices with an adequately powered sample using measures of known reliability. Prior studies have been inconclusive because it has been difficult to gather large amounts of data to establish test-retest reliability. For instance, van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks. However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of the individual measures. In the pre-registration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (model A) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model B), or a base model where each task is independently lateralised (model C). Given our previous findings using fTCD to study dissociations in language laterality (Woodhead, Bradshaw, Wilson, Thompson, & Bishop, 2019), we further predicted that model fit may differ for right- and left-handers, with more dissociation between tasks in the latter group.

@@

In practice, our final sample exceeded 300, because we discarded participants from individual tasks for failing to meet prespecified criteria (see below). In total, 441 native English speakers were initially recruited through Prolific Academic (www.prolific.co)`r colFmt(", with 411 participants completing both sessions",'red')`. Participants `r colFmt("were 18- to 50-years-old,",'red')`, had normal or corrected-to-normal vision and reported no history of neurological disease (e.g. head injury, brain tumour, stroke) or significant hearing loss or impairment.

@@

Adapted versions of the Miles test (Miles, 1929) and the Porta test (Porac & Coren, 1976) were used to determine each participant’s eye dominance in central gaze (i.e., when looking straight ahead). Each test classified participants as being either left or right eye dominant. `r colFmt("While we do not use tests of ocular dominance to asnwer apriori predictions, we report eyedness for participants inorder to further characterise them.",'red')` In practice, the two tests agreed perfectly, so we report here just the Miles test.

@@

The rhyme decision task was developed for this project as a novel visual half-field technique designed to engage lateralised language processes. On each trial, participants were asked to decide whether a foveally presented letter string rhymed with the name of an image shown in either the LVF or RVF. With online presentation, a particular concern is a potential for eye movements away from fixation. We relied on the fact that participants had to focus initially on the central written word to do the task `r colFmt(". We believed that this would encourage fixation to a central location at the start of the trial. We subsequently displayed pictures under",'red')`, brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982). Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972; but see also Bourne, 2006). Pictures of familiar objects were presented for rhyme matching to the written word, to ensure that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing. `r colFmt("While previous studies (e.g. Hunter",'red')` & `r colFmt("Brysbaert, 2008) have used overt naming, doing so online presents several problems. First, accurate naming latencies would require a microphone. This would have to be of sufficent quality to accurate capture naming latencies which many participants may not have. Seond, the recording of auditory responses would require additional scoring following data collection which would be a taxing procedure for researchers. It would also require additional ethical considerations with regards to the storing and processing of the data. It is for these reasons that we adopted a covert naming approach.",'red')`

@@

![Schematic illustration of a trial on the rhyme decision task. In the example, the word “bite” is presented along with the two images “kite” and “door”. When making the correct judgement, participants would respond with by pressing ‘S’ to indicate that the rhyme was on the left.](Picture1.png){width=60%}

@@

Each trial began with a fixation cross presented centrally for 800 ms. Participants were instructed to keep their eyes on the central fixation point. Then a foveally presented word would appear for 200 ms. The bilaterally presented images were then presented for 150 ms. At stimulus offset, participants indicated whether the name of the image in the LVF or RVF rhymed with the foveally presented word. Participants were instructed to press ‘S’ if the written word rhymed with the image in the LVF, and ‘K’ if it rhymed with the image in the RVF. If participants did not detect a rhyme, they were instructed to press the space bar. The response triggered the next trial`r colFmt("; trials did not time-out after a predetermined duration",'red')`. Accuracy and reaction times (RT) were recorded.

@@

As the dichotic listening task demands the use of stereo-headphones, participants were screened at the start of each session. Screening consisted of two tasks. The first involved participants judging which of three pure tones was quietest, and is described by Woods, Siegel, Traer, and McDermott (2017) in full detail. This sound discrimination task was designed to be easy to achieve when using headphones but difficult over loudspeakers due to phase-cancellation. Participants were removed from the study if they failed to correctly discriminate the quietest sound on at least five out of six trials, indicating that they were unlikely to be using headphones. The second check involved participants hearing a sound presented in only one of the two channels. Participants were removed if they failed to correctly identify which channel the sound was played on five out of six trials, indicating that they were not using stereo-headphones [^1].

Participants completed four blocks of the dichotic listening task. In each block, participants heard each of the 36 stimuli once. Across the four blocks, they responded to 144 stimuli. At the start of each trial, participants saw a fixation cross for 250 ms, after which a stimulus was played. After hearing each stimulus, participants were instructed to report the syllable they heard or if it seemed like they heard two different sounds, the one they heard most clearly by clicking a button corresponding to one of the six syllables. The response triggered the start of the next trial`r colFmt("; trials did not time-out after a predetermined duration",'red')`. Accuracy and reaction times (RT) were recorded.

@@

\

![An example stimulus from the chimeric faces task. In the example, “happiness” is shown on the left and “sadness” on the right. A participant who is right lateralised for emotional processing would be more likely to perceive the emotion in the left hemifield and would, therefore, respond “happiness”.](chim_example.png){width=20%}

\

@@

Participants completed four blocks of the chimeric faces task. In each block, participants viewed each stimulus twice. Thus, they viewed 32 stimuli per block and 128 in total. At the start of each trial, participants fixated a cross for 1000 ms. A chimeric face was then displayed for 400 ms. Participants then reported the emotion seen in the face by clicking a button corresponding to one of the four emotions. The response triggered the start of the next trial`r colFmt("; trials did not time-out after a predetermined duration",'red')`. Accuracy and reaction times (RT) were recorded.

@@

The current study implemented a test-retest, within-subjects design across two sessions. Each session was three to 10 days apart. During session 1, participants completed the full battery of tasks. Self -report measures were completed first, followed by the behavioural laterality tasks. During session 2, participants completed only the behavioural laterality tasks. Hence, each participant provided data for the rhyme decision, dichotic listening, chimeric faces, and finger tapping tasks twice. Each session lasted approximately 45 minutes and 35 minutes respectively. `r colFmt("Participants were paid £15 for their participation.",'red')`

When completing the behavioural tasks participants were instructed to sit approximately 50 cm from the screen. Gorilla’s scaling tool was utilised such that stimuli size and eccentricity was maintained across tasks and participants. `r colFmt("During each task participants were instructed respond as fast and accurately as possible.",'red')` Stimuli from each task were repeated across four blocks. Within each block, participants completed 70 trials of the rhyme detection task (R); 36 trials of the dichotic listening task (D); 32 trials of the chimeric face task (C); and 2 trials of the finger tapping task (F) for both their dominant and non-dominant hand. The order of tasks with each block was determined as follows RDCF, DCFR, CFRD, FRDC, and the order of tasks remained consistent across blocks for each participant. For each task, trial order was randomised, except for the finger tapping task where participants responded with their dominant hand followed by their non-dominant hand.

@@

For each task, we adopted the following data cleaning procedures. First, for all tasks, we removed trials in which response times were shorter than 200 ms. Response times were log-transformed to reduce skew, as specified in our pre-registration. Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)) [^2]. Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution. The Hoaglin and Iglewicz procedure was applied to each participant's data separately. After applying each data trimming procedure, participants left with less than 80% of trials were removed.

@@

Foot and eye preference characteristics of the sample after exclusions as noted above are shown in Table 1. As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we report data only from the Miles test. `r colFmt("As eye dominance is not a focus of this study, we will not elaborate further on it.", 'red')`

@@

}
```
\

Table 2. Mean LIs and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality for the rhyme decision (RD), dichotic listening (DL), chimeric faces (CF), and finger tapping (FT) tasks at sessions 1 and 2.\
```{r Tab2,echo=FALSE, comment=FALSE, warning=FALSE}
tab2 <- flextable(LItable)
tab2 <- align(tab2, j= 2:7, align= "center")
autofit(tab2, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
```

*Note*. LIs are scored such that a positive value equates to left lateralisation at the brain level.

\
